String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-08-18 2023-1-19/sections/abstract.tex
some little errors

%\documentclass{standalone}
%\begin{document}

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. 
%Solving string constraints involving regular membership and string-length function has been widely investigated. 
Counting is a frequently used feature in regexes that counts the number of sub-patterns matchings. The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length efficiently, especially when the counting bounds are large. In this work, we propose an automata-theoretic approach for solving such a class of string constraints. 
%
The main idea is to model the counting operators symbolically by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem.  
%
Moreover, the string-length function is modeled by register as well. 
%Since Parikh images of automata can be defined by linear integer arithmetic formulas, 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which off-the-shelf SMT solvers can then solve. 
%
%we extract linear integer arithmetic formulas from automata, which encode the constraints on the values of registers, and utilize the off-the-shelf SMT solvers to solve the linear integer arithmetic constraints. 
%
To improve the performance further, we propose techniques to reduce automata sizes.
% we propose state reduction techniques and under-approximation to improve the performance further. 
We implement the algorithms and validate our approach on 48,843 benchmark instances. The experimental results show that 
our approach can solve more instances than other solvers at a comparable or faster speed.
% performance of this approach compared against the state-of-the-art string solvers.

%Regular expression with repetition frequently appears in the real world, such as $(ab)\{1, 100\}$. However, state-of-art string
%solvers can not efficiently solve string constraints with repetition bounds,
%especially when the repetition bounds are large. This paper focuses on
%string logic containing regular membership predicate with repetition
%and linear integer constraints on string length. Repetition is captured
%by cost-enriched finite automaton(CEFA), whose transitions have symbolic updates of integers limited by linear arithmetic. We propose a new systematic algorithm based on CEFA, and heuristic ways
%like under-approximation and symbolic-aware simplification are used to
%accelerate. We implement a powerful string solver OstrichCEA and
%evaluate it on instances from the real world and other typical benchmarks.
%The experiment result shows superiority.

%\keywords{Rgular Expressions, Counting, Length, Automata, Registers, Linear Integer Arithmetic}
%\end{document}

>>>>

Regular expressions (regex for short) and string-length function are widely used in string-manipulating programs. Counting is a frequently used feature in regexes that counts the number of matchings of sub-patterns. 
The state-of-the-art string solvers are incapable of solving string constraints with regex-counting and string-length eﬀiciently, especially when the counting and length bounds are large. 
In this work, we propose an automata-theoretic approach for solving such a class of string constraints. 
The main idea is to model the counting operators symbolically by registers in automata instead of unfolding them explicitly, thus alleviating the state explosion problem. Moreover, the string-length function is modeled by a register as well. 
As a result, the satisfiability of string constraints with regex-counting and string-length is reduced to the satisfiability of linear integer arithmetic, which the off-the-shelf SMT solvers can then solve. 
To improve the performance further, we propose techniques to reduce automata sizes. We implement the algorithms and validate our approach on 48,843 benchmark instances. 
The experimental results show that our approach can solve more instances than the other solvers, at a comparable or faster speed, especially when the counting and length bounds are large.

### 2023-08-18 2023-1-19/sections/introduction.tex
some little errors

Nevertheless, the aforementioned state-of-the-art string constraint solvers still suffer from such string constraints, especially when the counting bounds are large. For instance, none of the string solvers CVC5, Z3seq, Z3-Trau, Z3str3, Z3str3RE, and OSTRICH is capable of solving the following constraint within 300 seconds, and Z3 can not solve it within 120 seconds.
%
\begin{equation}\label{eqn-running}
x \in (\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}} \wedge x \in \Sigma^* c^+ \wedge |x| > 120.
\end{equation}
Intuitively, the constraint in (\ref{eqn-running}) specifies that $x$ is a concatenation of three strings $x_1, x_2, x_3$ where $a$ (resp. $b, c$) does not occur in $x_1$ (resp. $x_2, x_3$), moreover, $x$ ends with a nonempty sequence of $c$'s, and the length of $x$ is greater than $120$. It is easy to observe that this constraint is unsatisfiable since on the one hand, $|x| > 120$ and the counting upper bound $60$ in both $(\Sigma \setminus a)^{\{1, 60\}}$ and $(\Sigma \setminus b)^{\{1, 60\}}$ imply that $x$ must end with some character from $\Sigma \setminus c$, that is, a character different from $c$, and on the other hand, $x \in \Sigma^*c^+$ requires that $x$ has to end with $c$.

>>>>

Nevertheless, the aforementioned state-of-the-art string constraint solvers still suffer from such string constraints, especially when the counting bounds and length bounds are large. For instance, none of the string solvers CVC5, Z3seq, Z3-Trau, Z3str3, Z3str3RE, and OSTRICH is capable of solving the following constraint within 300 seconds, and Z3 can not solve it within 120 seconds.
%
\begin{equation}\label{eqn-running}
x \in (\Sigma \setminus a)^{\{1, 60\}} (\Sigma \setminus b)^{\{1, 60\}} (\Sigma \setminus c)^{\{0, 60\}} \wedge x \in \Sigma^* c^+ \wedge |x| > 120.
\end{equation}
Intuitively, the constraint in (\ref{eqn-running}) specifies that $x$ is a concatenation of three strings $x_1, x_2, x_3$ where $a$ (resp. $b, c$) does not occur in $x_1$ (resp. $x_2, x_3$), moreover, $x$ ends with a nonempty sequence of $c$'s, and the length of $x$ is greater than $120$. It is easy to observe that this constraint is unsatisfiable since on the one hand, $|x| > 120$ and the counting upper bound $60$ in both $(\Sigma \setminus a)^{\{1, 60\}}$ and $(\Sigma \setminus b)^{\{1, 60\}}$ imply that $x$ must end with some character from $\Sigma \setminus c$, that is, a character different from $c$, and on the other hand, $x \in \Sigma^*c^+$ requires that $x$ has to end with $c$.

### 2023-08-18 2023-1-19/sections/overview.tex
some little errors

Specifically, to reduce the sizes of the $\cB$, we determinize $\cB$ and apply the minimization algorithm to the resulting deterministic finite automaton (DFA), obtaining a DFA $\cC$, as illustrated in Figure~\ref{fig:overview:product:reduced}(b).
Note that $\cC$ contains only three states $q''_0, q''_1, q''_2$ and six transitions, while $\cB$ contains four states and eight transitions. 
The evaluation in section \cite{sec} shows that we can solve 1503 more instances after utilizing simplification techniques, with \%54.8 less average time.

>>>>

Specifically, to reduce the sizes of the $\cB$, we determinize $\cB$ and apply the minimization algorithm to the resulting deterministic finite automaton (DFA), obtaining a DFA $\cC$, as illustrated in Figure~\ref{fig:overview:product:reduced}(b).
Note that $\cC$ contains only three states $q''_0, q''_1, q''_2$ and six transitions, while $\cB$ contains four states and eight transitions. 
The evaluation in Table~\ref{tab:results_simp} shows that we can solve 1503 more instances after utilizing simplification techniques, with \%54.8 less average time.
